Wigglyhedra

### 2024-10-11 FPSAC25/wigglyhedraFPSAC.tex
start FPSAC

\documentclass{amsart}

\usepackage[T1]{fontenc}
\usepackage{enumerate, amsmath, amsfonts, amssymb, amsthm, mathrsfs, wasysym, graphics, graphicx, xcolor, url, hyperref, hypcap, shuffle, xargs, multicol, overpic, pdflscape, multirow, hvfloat, minibox, accents, array, multido, xifthen, a4wide, ae, aecompl, blkarray, pifont, mathtools, etoolbox, dsfont}
\usepackage{marginnote}
\hypersetup{colorlinks=true, citecolor=darkblue, linkcolor=darkblue}
\usepackage[all]{xy}
\usepackage[bottom]{footmisc}
\usepackage{tikz}
%\usepackage{tkz-graph}
%\usepackage{tikz-qtree}
\usetikzlibrary{trees, decorations, decorations.markings, shapes, arrows, matrix, calc, fit, intersections, patterns, angles, cd}
\usepackage[external]{forest}
%\tikzexternalize
\graphicspath{{figures/}}
\makeatletter\def\input@path{{figures/}}\makeatother
\usepackage{caption}
\captionsetup{width=\textwidth}
\renewcommand{\topfraction}{1} % possibility to have one page of pictures
\renewcommand{\bottomfraction}{1} % possibility to have one page of pictures
\usepackage[noabbrev,capitalise]{cleveref}
\usepackage[export]{adjustbox}
\usepackage{ulem}\normalem

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% formating the table of contents
\setcounter{tocdepth}{4}
\makeatletter
\def\l@part{\@tocline{1}{8pt}{0pc}{}{}}
\def\l@section{\@tocline{1}{4pt}{0pc}{}{}}
\makeatother
\let\oldtocpart=\tocpart
\renewcommand{\tocpart}[2]{\sc\large\oldtocpart{#1}{#2}}
\let\oldtocsection=\tocsection
\renewcommand{\tocsection}[2]{\bf\oldtocsection{#1}{#2}}
\let\oldtocsubsubsection=\tocsubsubsection
\renewcommand{\tocsubsubsection}[2]{\quad\oldtocsubsubsection{#1}{#2}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Wigglyhedra}

\thanks{AB was partially supported by the Australian Research Council DECRA grant DE240100447. 
  VP was partially supported by the Spanish grant PID2022-137283NB-C21 of MCIN/AEI/10.13039/501100011033 / FEDER, UE, by Departament de Recerca i Universitats de la Generalitat de Catalunya (2021 SGR 00697), by the French grant CHARMS (ANR-19-CE40-0017), and by the French--Austrian projects PAGCAP (ANR-21-CE48-0020 \& FWF~I~5788).}

\author{Asilata Bapat}
\address{The Australian National University}
\email{asilata.bapat@anu.edu.au}
\urladdr{\url{https://asilata.github.io}}

\author{Vincent Pilaud}
\address{Universitat de Barcelona}
\email{vincent.pilaud@ub.edu}
\urladdr{\url{https://www.ub.edu/comb/vincentpilaud/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{abstract}
Motivated by categorical representation theory, we define the wiggly complex, whose vertices are arcs wiggling around $n+2$ points on a line, and whose faces are sets of wiggly arcs which are pairwise pointed and non-crossing. The wiggly complex is a $(2n-1)$-dimensional pseudomanifold, whose facets are wiggly pseudotriangulations. We show that wiggly pseudotriangulations are in bijection with wiggly permutations, which are permutations of~$[2n]$ avoiding the patterns~$(2j-1) \cdots i \cdots (2j)$ for~$i < 2j-1$ and~$(2j) \cdots k \cdots (2j-1)$ for~$k > 2j$. These permutations define the wiggly lattice, an induced sublattice of the weak order. We then prove that the wiggly complex is isomorphic to the boundary complex of the polar of the wigglyhedron, for which we give explicit and simple vertex and facet descriptions. Interestingly, we observe that any Cambrian associahedron is normally equivalent to a well-chosen face of the wigglyhedron. Finally, we recall the correspondence of wiggly arcs with objects in a category, and we develop categorical criteria for a subset of wiggly arcs to form a face of the wiggly complex.
\end{abstract}

\maketitle

\centerline{\includegraphics[scale=1.75]{wigglyFlipGraph}}

\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

In this paper we consider wiggly pseudotriangulations, which are built from arcs wiggling around points on a line.
We study their combinatorics in detail, construct polyhedral realizations of their flip graph, and interpret them in terms of categorical representation theory.

\bigskip
\para{Combinatorics}
Fix $n \ge 1$ and consider the~$n+2$ points on the horizontal axis labelled~${0, \dots, n+1}$.
A \defn{wiggly arc} is an abscissa monotone curve starting at some point~$i$, ending at some point~$j$ (with~${0 \le i < j \le n+1}$), and passing above or below each point~$i+1, \dots, j-1$.
Two wiggly arcs are \defn{compatible} if they are pointed (the left endpoint of one is not the right endpoint of the other) and non-crossing (the interiors of the curves are disjoint).
A \defn{wiggly dissection} is a collection of pairwise compatible wiggly arcs, and a \defn{wiggly pseudotriangulation} is an inclusion maximal wiggly dissection.
The \defn{wiggly complex}~$\wigglyComplex_n$ is the simplicial complex of wiggly dissections (minus their external wiggly arcs).
We first establish the following result.

\begin{proposition*}
The wiggly complex~$\wigglyComplex_n$ is a $(2n-1)$-dimensional pseudomanifold without boundary.
\end{proposition*}

In particular, all wiggly pseudotriangulations have $2n-1$ internal wiggly arcs, each of which can be \defn{flipped}.
We actually describe the flip in terms of wiggly diagonals in wiggly pseudoquadrangles.
We also orient this flip graph in a canonical way, which we call the \defn{wiggly increasing flip graph}~$\wigglyIncreasingFlipGraph_n$.
The wiggly increasing flip graph on wiggly pseudotriangulations of $4$ points is the front page illustration of this paper (throughout the paper, we implicitly orient the graphs and posets from bottom to top).

We then make a small detour to permutations.
A \defn{wiggly permutation} is a permutation of~$[2n]$ avoiding the patterns~$(2j-1) \cdots i \cdots (2j)$ for~$i < 2j-1$ and~$(2j) \cdots k \cdots (2j-1)$ for~$k > 2j$.
Our main result on wiggly permutations is the following.

\begin{proposition*}
The wiggly permutations induce a sublattice of the weak order on permutations of~$[2n]$, which we call the \defn{wiggly lattice}~$\wigglyLattice_n$. Moreover, its Hasse diagram is regular of degree~$2n-1$.
\end{proposition*}

We then connect wiggly pseudotriangulations to wiggly permutations.

\begin{proposition*}
The increasing flip graph on wiggly pseudotriangulations~$\wigglyIncreasingFlipGraph_n$ is isomorphic to the Hasse diagram of the wiggly lattice~$\wigglyLattice_n$. In particular, the wiggly pseudotriangulations admit a lattice structure.
\end{proposition*}

\para{Polyhedral geometry}
We then switch to geometric realizations of the wiggly complex~$\wigglyComplex_n$.
Our constructions are inspired by similar constructions for generalized associahedra~\cite{HohlwegLangeThomas,HohlwegPilaudStella} of finite type cluster algebras~\cite{FominZelevinsky-ClusterAlgebrasI,FominZelevinsky-ClusterAlgebrasII}, and for gentle associahedra~\cite{PaluPilaudPlamondon-nonkissing} of support \mbox{$\tau$-tilting} complexes~\cite{AdachiIyamaReiten} on gentle algebras~\cite{ButlerRingel}.
Namely, we first construct a fan supported by certain $\b{g}$-vectors associated to the wiggly arcs.

\begin{theorem*}
The cones generated by the $\b{g}$-vectors of wiggly pseudodissections form a complete simplicial fan, called \defn{wiggly fan}~$\wigglyFan_n$, realizing the wiggly complex~$\wigglyComplex_n$.
\end{theorem*}

The proof of this result is based on a fine understanding of the linear dependences between \mbox{$\b{g}$-vectors}.
We further use this understanding to find a height function~$\kappa$ on wiggly arcs which satisfies all wall-crossing inequalities of the wiggly fan~$\wigglyFan_n$.
Without entering the details of the definition of~$\kappa$ at the moment, let us just mention that it is obtained by summing some compatibility degrees of wiggly arcs, analogously to the constructions for cluster algebras~\cite{HohlwegPilaudStella} and gentle algebras~\cite{PaluPilaudPlamondon-nonkissing}.
Using the $\b{c}$-vectors of each pseudotriangulation, defined to form the dual basis of the $\b{g}$-vectors, we obtain our main result.

\begin{theorem*}
The wiggly fan~$\wigglyFan_n$ is the normal fan of a simplicial $(2n-1)$-dimensional polytope, called the \defn{wigglyhedron}~$\wigglyhedron_n$, and defined equivalently~as
\begin{itemize}
\item the intersection of the halfspaces~$\set{\b{x} \!\in\! \HH_{2n}\!}{\!\dotprod{\b{g}(\alpha)}{\b{x}} \!\le\! \kappa(\alpha)}$ for all internal wiggly arcs~$\alpha$,
\item the convex hull of the points~$\b{p}(T) \eqdef \sum\limits_{\alpha \in T} \kappa(\alpha) \, \b{c}(\alpha, T)$ for all wiggly pseudotriangulations~$T$.
\end{itemize}
\end{theorem*}

Moreover, as for the classical permutahedron and associahedron, we recover our lattice structure by a suitable linear orientation of the graph of the wigglyhedron.

\begin{proposition*}
The Hasse diagram of the wiggly lattice~$\wigglyLattice_n$ (or equivalently, the wiggly increasing flip graph~$\wigglyIncreasingFlipGraph_n$) is isomorphic to the graph of the wigglyhedron~$\wigglyhedron_n$ oriented in a suitable direction.
\end{proposition*}

\para{Further topics}
The wigglyhedron has interesting connections in lattice theory and discrete geometry.
We first connect the wigglyhedron to the (type~$A$) Cambrian world.

\begin{theorem*}
For any~$\delta \in \{-,+\}^n$,
\begin{itemize}
\item the $\delta$-Cambrian lattice of~\cite{Reading-CambrianLattices} is an interval in the wiggly lattice~$\wigglyLattice_n$,
\item the $\delta$-Cambrian fan of~\cite{ReadingSpeyer} is a link in the wiggly fan~$\wigglyFan_n$,
\item the $\delta$-associahedron of~\cite{HohlwegLange} is normally equivalent to a face of the wigglyhedron~$\wigglyhedron_n$.
\end{itemize}
\end{theorem*}

We then define the wiggly complex~$\wigglyComplex_P$ of an arbitrary planar point set~$P$ (not necessarily aligned along the horizontal axis, nor necessarily in general position).
Our definition recovers:
\begin{itemize}
\item when $P$ consists of $n+2$ collinear points, the wiggly complex~$\wigglyComplex_n$ considered in this~paper,
\item when~$P$ is in general position, the pseudodissection complex of~$P$ considered in discrete and computational geometry~\cite{PocchiolaVegter,RoteSantosStreinu-pseudotriangulations}, and realized by the pseudotriangulation polytope in~\cite{RoteSantosStreinu-polytope} using rigidity and expansive motions.
\end{itemize}
These two families of examples motivate the following conjecture.

\begin{conjecture*}
For any point set~$P$ in the plane, the wiggly complex~$\wigglyComplex_P$ is the boundary complex of a simplicial polytope.
\end{conjecture*}

We also present a few open questions related to this conjecture, concerning in particular the possible interpretation of the wiggly complex~$\wigglyComplex_P$ in terms of dual pseudoline arrangements, and the extension to multi wiggly complexes and arbitrary finite Coxeter groups.

\pagebreak
\para{Categorical interpretation}
Each wiggly arc on \(n+2\) points corresponds to an object of an abelian category, such that morphisms and extensions between objects in the category are captured by minimal intersection numbers between the wiggly arcs.
In fact, this correspondence holds for more general curves, which then correspond to objects of a larger triangulated category~\cite{kho.sei:02}.

A general curve can be decomposed into wiggly arcs simply by cutting along its U-turns.
We interpret this operation categorically as the decomposition of an object of the triangulated category into its cohomology pieces, from which we deduce the following.

\begin{proposition*}
The set of wiggly arcs corresponding to the cohomology pieces of an object represented by a general arc is a face of the wiggly complex.
\end{proposition*}

Furthermore, we characterize categorically when two objects in the abelian category are compatible, that is, their corresponding wiggly arcs are compatible.

\begin{proposition*}
Two wiggly arcs are compatible if and only if there are no extensions between the corresponding objects in the abelian category.
\end{proposition*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Wiggly pseudotriangulations and wiggly permutations}
\label{sec:combinatorics}

Fix an integer~$n \ge 1$.
In this section, we define two combinatorial families, the wiggly pseudotriangulations of~$n+2$ points (\cref{subsec:wigglyPseudotriangulations}) and the wiggly permutations of~$[2n]$ (\cref{subsec:wigglyPermutations}), and we prove that they are in bijection (\cref{subsec:bijection}).
These two perspectives enable us to endow these combinatorial families with additional structures: the wiggly pseudotriangulations are actually the facets of the wiggly complex, while the wiggly permutations are actually the elements of the wiggly lattice.
We conclude with a few questions and conjectures on graph properties of the wiggly flip graph (\cref{subsec:graphProperties}).

%%%%%%%%%%%%

\subsection{Wiggly pseudotriangulations and the wiggly complex}
\label{subsec:wigglyPseudotriangulations}

We start with wiggly pseudodissections of $n+2$ points on a line.
The following two definitions are illustrated in \cref{fig:incompatible}.

\begin{definition}
A \defn{wiggly arc} is a quadruple $\alpha \eqdef (i,j,A,B)$ where $0 \le i < j \le n+1$ and the sets~$A$ and~$B$ form a partition of~${]i,j[} \eqdef \{i+1, \dots, j-1\}$.
We represent $\alpha$ by an $x$-monotone curve wiggling around the horizontal axis, starting at point~$i$, ending at point~$j$, and passing above the points of~$A$ and below the points of~$B$.
The wiggly arcs~$\alpha_\mathrm{top} \eqdef (0, n+1, [n], \varnothing)$ and~$\alpha_\mathrm{bot} \eqdef (0, n+1, \varnothing, [n])$ are called \defn{external}, all other wiggly arcs are called \defn{internal}.
\end{definition}

\begin{remark}
\label{rem:numberWigglyArcs}
The number of wiggly arcs is~$\sum_{\ell = 0}^n 2^\ell (n+1-\ell)$.
\end{remark}

\begin{definition}
\label{def:compatible}
Two wiggly arcs~$(i,j,A,B)$ and~$(i',j',A',B')$ are 
\begin{itemize}
\item \defn{non pointed} if~$i = j'$ or~$i' = j$,
\item \defn{crossing} if~$(A \cap B') \cup (\{i,j\} \cap B') \cup (A \cap \{i',j'\}) \ne \varnothing \ne (A' \cap B) \cup (\{i',j'\} \cap B) \cup (A' \cap \{i,j\})$, that is, if the corresponding curves cross,
\item \defn{incompatible} if they are non pointed or crossing, and \defn{compatible} otherwise.
\end{itemize}
%
\begin{figure}[b]
\centerline{\includegraphics[scale=1.3]{incompatible}}
\caption{Some incompatible wiggly arcs: non pointed (left) and crossing (right).}
\label{fig:incompatible}
\end{figure}
\end{definition}

\begin{remark}
We want to stress out that our notion of compatibility on wiggly arcs of \cref{def:compatible} might seem similar to, but fundamentally differs from the notion of compatibility used by N.~Reading to define noncrossing arc diagrams~\cite{Reading-arcDiagrams}.
Namely, our wiggly arcs and our crossing condition coincide with that of~\cite{Reading-arcDiagrams}.
However, we forbid an arc to start where another arc ends but allow two arcs to start or end at the same point,
%for two wiggly arcs~$\alpha$ and~$\alpha'$ to be compatible, we forbid the left endpoint of~$\alpha$ to coincide with the right endpoint of~$\alpha'$, but allow $\alpha$ and~$\alpha'$ to have the same left endpoints or the same right endpoints, 
while N.~Reading does precisely the opposite in~\cite{Reading-arcDiagrams}.
%This results in very different objects.
\end{remark}

We then consider pairwise compatible sets of wiggly arcs.
The following two definitions are illustrated~in~\cref{fig:pseudodissections}.

\begin{definition}
A \defn{wiggly pseudodissection}~$D$ is a set of pairwise compatible wiggly arcs which contains the exterior wiggly arcs~$\alpha_\mathrm{top}$ and~$\alpha_\mathrm{bot}$. We denote by~$D^\circ \eqdef D \ssm \{\alpha_\mathrm{top}, \alpha_\mathrm{bot}\}$.
%
\begin{figure}
\centerline{\includegraphics[scale=1.5]{wigglyPseudodissections}}
\caption{Three wiggly pseudodissections for~$n = 7$. The left one has three wiggly cells, of respective degrees $3$ (red), $8$ (green), and $4$ (blue). The other two have $7$ pseudotriangles, and are obtained from each other by flipping the wiggly arc separating their yellow and dark blue pseudotriangles.}
\label{fig:pseudodissections}
\end{figure}
\end{definition}

%\vincent{Here, I am using the fact that a set of pairwise non-crossing wiggly arcs can be drawn simultaneously non-crossing. I am not sure I want to enter these details.}

\begin{definition}
A wiggly pseudodissection~$D$ decomposes the digon bounded by~$\alpha_\mathrm{top}$ and~$\alpha_\mathrm{bot}$ into cells, that we call \defn{wiggly cells}.
The \defn{degree}~$\delta_c$ of a wiggly cell~$c$ is~$\delta_c \eqdef \beta_c/2+2\kappa_c-1$ where~$\beta_c$ is the number of wiggly arcs on the boundary of~$c$ (counted twice if they appear twice along the boundary of~$c$), and~$\kappa_c$ is the number of connected components of the boundary of~$c$.
Two consecutive wiggly arcs along the boundary of~$c$ define a \defn{corner} (resp.~a \defn{hinge}) of~$c$ if they bound a convex (resp.~concave) angle of~$c$.
%The \defn{concave chains} of~$c$ are the maximal sequences of wiggly arcs on the external boundary of~$c$ connected by hinges of~$c$.
\end{definition}

\begin{remark}
\label{rem:degree}
Consider a wiggly cell~$c$ in a wiggly pseudodissection~$D$.
Note that~$\beta_c$ is always even as~$D$ is pointed, so that~$\beta_c/2$ is an integer.
Moreover, we have~$\beta_c \ge 2$ and~$\kappa_c \ge 1$, where at least one of the two inequalities is strict (as otherwise, the two wiggly arcs bounding~$c$ would coincide).
%We conclude that~$\delta_c = \beta_c/2+2\kappa_c-1 \ge 3$ with equality if and only if~$\beta_c = 4$ and~$\kappa_c = 1$.
We thus obtain on the one hand that~$\delta_c = \beta_c/2+2\kappa_c-1 \ge 3$ with equality if and only if~$\beta_c = 4$ and~$\kappa_c = 1$, and on the other hand that~$\beta_c/2+\kappa_c-1 \ge 2$.
\end{remark}

\begin{definition}
A \defn{wiggly pseudotriangle} (resp.~\defn{pseudoquadrangle}) is a wiggly cell of degree~$3$ (resp.~$4$).
\end{definition}

>>>>

%% if you are submitting an initial manuscript then you should have submission as an option here
%% if you are submitting a revised manuscript then you should have revision as an option here
%% otherwise options taken by the article class will be accepted
\documentclass[submission]{FPSAC2025}
%% but DO NOT pass any options (or change anything else anywhere) which alters page size / layout / font size etc

%% note that the class file already loads {amsmath, amsthm, amssymb}

\usepackage{xargs}
\usepackage{hypcap}
\usepackage{paralist}
\usepackage{dsfont}
\usepackage{mathtools}
\usepackage{blkarray}
\usepackage{picins/picins}
\usepackage{overpic}

@@

\graphicspath{{figures/}}

%% define your title in the usual way
\title{Wigglyhedra}

%% define your authors in the usual way
%% use \addressmark{1}, \addressmark{2} etc for the institutions, and use \thanks{} for contact details
%% for more than 3 names, use the Oxford comma: One Author, Two Author, Red Author, \and Blue Author
\author{Asilata Bapat\thanks{\href{mailto:asilata.bapat@anu.edu.au}{mailto:asilata.bapat@anu.edu.au}. Supported by the Australian DECRA grant DE240100447.}\addressmark{1} \and Vincent Pilaud\thanks{\href{mailto:vincent.pilaud@ub.edu}{mailto:vincent.pilaud@ub.edu}. Supported by the French--Austrian grant PAGCAP (ANR-21-CE48-0020 \& FWF I 5788) and by the Spanish grant PID2022-137283NB-C21.}\addressmark{2}}

%% then use \addressmark to match authors to institutions here
\address{\addressmark{1}The Australian National University, Canberra, Australia \\ \addressmark{2}Universitat de Barcelona, Spain}

%% put the date of submission here
\received{\today}

%% leave this blank until submitting a revised version
%\revised{}

%% put your English abstract here, or comment this out if you don't have one yet
%% please don't use custom commands in your abstract / resume, as these will be displayed online
%% likewise for citations -- please don't use \cite, and instead write out your citation as something like (author year)
\abstract{Motivated by categorical representation theory, we study the wigglyhedron. Its boundary complex encodes wiggly dissections (sets of pairwise pointed and non-crossing arcs wiggling around $n+2$ points on a line). Its graph is the Hasse diagram of the sublattice of the weak order induced by wiggly permutations (permutations of $[2n]$ avoiding~$2j-1 \cdots i \cdots 2j$ for~$i < 2j-1$ and~$2j \cdots k \cdots 2j-1$ for~$k > 2j$). It contains copies of all type~$A_n$ Cambrian lattices and Cambrian fans.}

%% put your French abstract here, or comment this out if you don't have one
\resume{Motivés par la théorie catégorique des représentations, nous étudions le slalomèdre. Son complexe de bord encode les dissections slalomantes (ensembles deux-à-deux pointés et sans croisement d'arcs slalomant autour de $n+2$ points sur une ligne). Son graphe est le diagramme de Hasse du sous-treillis de l'ordre faible induit par les permutations slalomantes (permutations de $[2n]$ évitant~$2j-1 \cdots i \cdots 2j$ pour~$i < 2j-1$ et~$2j \cdots k \cdots 2j-1$ pour~$k > 2j$). Il contient des copies de tous les treillis Cambriens et associaèdres Cambriens de type~$A_n$.}

%% put your keywords here, or comment this out if you don't have them yet
%\keywords{math, maths, mathematics}

%% you can include your bibliography however you want, but using an external .bib file is STRONGLY RECOMMENDED and will make the editor's life much easier
%% regardless of how you do it, please use numerical citations; i.e., [xx, yy] in the text

%% this sample uses biblatex, which (among other things) takes care of URLs in a more flexible way than bibtex
%% but you can use bibtex if you want
\usepackage[backend=bibtex]{biblatex}
\addbibresource{wigglyhedra.bib}
%% note the \printbibliography command at the end of the file which goes with these biblatex commands

@@

\stepcounter{footnote}
\stepcounter{footnote}

\maketitle
%% note that you DO NOT have to put your abstract here -- it is generated by \maketitle and the \abstract and \resume commands above

Motivated by categorical representation theory, we consider the wiggly complex (\cref{sec:wigglyPseudotriangulations}), whose vertices are arcs wiggling around $n+2$ points on a line, and whose faces are sets of pairwise pointed and non-crossing wiggly arcs.
It is a $(2n-1)$-dimensional manifold whose maximal faces are wiggly pseudotriangulations of a disc.
We show (\cref{sec:bijection}) that the flip graph on wiggly pseudotriangulations is isomorphic to the cover graph of the sublattice of the weak order induced by wiggly permutations (\cref{sec:wigglyPermutations}), the permutations of $[2n]$ avoiding~$2j-1 \cdots i \cdots 2j$ for~$i < 2j-1$ and~$2j \cdots k \cdots 2j-1$ for~$k > 2j$.
Inspired by similar constructions for finite type cluster complexes~\cite{HohlwegLangeThomas,HohlwegPilaudStella}, and for support $\tau$-tilting complexes on gentle algebras~\cite{PaluPilaudPlamondon-nonkissing}, we then provide geometric realizations of the wiggly complex.
We first construct the wiggly fan (\cref{sec:wigglyFan}), a complete simplicial fan realizing the wiggly complex and supported by the $\b{g}$-vectors of the wiggly arcs.
We then define the wigglyhedron (\cref{sec:wigglyhedron}), a polytope whose normal fan is the wiggly fan and whose support function is given by the incompatibility numbers of wiggly arcs.
Its graph can be oriented to obtain the Hasse diagram of the wiggly lattice, and it contains copies of all type~$A_n$ Cambrian lattice as faces (\cref{sec:CambrianConsiderations}).
To conclude, we conjecture that the wiggly complex of any planar point set is polytopal (\cref{sec:planarPointSets}), and we explore representation theoretic aspects of the wiggly complex (\cref{sec:representationTheory}).

\newpage
\section{Wiggly pseudotriangulations and the wiggly complex}
\label{sec:wigglyPseudotriangulations}

We start by defining the wiggly complex, the clique complex of a compatibility relation on the following wiggly arcs.

\begin{definition}
A \defn{wiggly arc} is a quadruple $\alpha \eqdef (i,j,A,B)$ where $0 \!\le\! i \!<\! j \!\le\! n+1$ and the sets~$A$ and~$B$ form a partition of~${]i,j[} \eqdef \{i+1, \dots, j-1\}$.
We represent $\alpha$ by an $x$-monotone curve wiggling around the horizontal axis, starting at point~$i$, ending at point~$j$, and passing above the points of~$A$ and below the points of~$B$.
The wiggly arcs~$\alpha_\mathrm{top} \eqdef (0, n+1, [n], \varnothing)$ and~$\alpha_\mathrm{bot} \eqdef (0, n+1, \varnothing, [n])$ are called \defn{external}, all other wiggly arcs are called \defn{internal}.
\end{definition}

\begin{definition}
\label{def:compatible}
Two wiggly arcs~$(i,j,A,B)$ and~$(i',j',A',B')$ are 
\begin{compactitem}
\item \defn{non pointed} if~$i = j'$ or~$i' = j$,
\item \defn{crossing} if their curves cross, that is, if $(A \cap B') \cup (\{i,j\} \cap B') \cup (A \cap \{i',j'\}) \ne \varnothing$ and $(A' \cap B) \cup (\{i',j'\} \cap B) \cup (A' \cap \{i,j\})\ne \varnothing$,
\item \defn{incompatible} if they are non pointed or crossing, and \defn{compatible} otherwise.
\end{compactitem}
%
%\begin{figure}[b]
%\centerline{\includegraphics[scale=1.3]{incompatible}}
%\caption{Some incompatible wiggly arcs: non pointed (left) and crossing (right).}
%\label{fig:incompatible}
%\end{figure}
\end{definition}

\begin{definition}
A \defn{wiggly pseudodissection}~$D$ is a set of pairwise compatible wiggly arcs which contains the exterior wiggly arcs~$\alpha_\mathrm{top}$ and~$\alpha_\mathrm{bot}$. We denote by~$D^\circ \eqdef D \ssm \{\alpha_\mathrm{top}, \alpha_\mathrm{bot}\}$.

\begin{figure}[b]
\centerline{\includegraphics[scale=1.4]{wigglyPseudodissections}}
\caption{Three wiggly pseudodissections for~$n = 7$. The left one has three wiggly cells, of respective degrees $3$ (red), $8$ (green), and $4$ (blue). The other two have $7$ pseudotriangles, and are obtained from each other by a flip.}
\label{fig:pseudodissections}
\end{figure}
\end{definition}

\begin{definition}
\label{def:wigglyComplex}
The \defn{wiggly complex}~$\wigglyComplex_n$ is the simplicial complex of pairwise pointed and non-crossing subsets of internal wiggly arcs.
%In other words, it is the clique complex of the compatibility graph on internal wiggly arcs.

\begin{figure}
\centerline{\includegraphics[scale=1.1]{wigglyComplex}}
\caption{The wiggly complex~$\wigglyComplex_2$ (left) and the wiggly flip graph~$\wigglyFlipGraph_2$ (right).}
\label{fig:wigglyComplex}
\end{figure}
\end{definition}

\begin{definition}
A wiggly pseudodissection~$D$ decomposes the digon bounded by~$\alpha_\mathrm{top}$ and~$\alpha_\mathrm{bot}$ into \defn{wiggly cells}.
The \defn{degree}~$\delta_c$ of a wiggly cell~$c$ is~$\delta_c \eqdef \beta_c/2+2\kappa_c-1$ where~$\beta_c$ is the number of wiggly arcs on the boundary of~$c$ (counted twice if they appear twice along the boundary of~$c$), and~$\kappa_c$ is the number of connected components of the boundary of~$c$.
Two consecutive wiggly arcs along the boundary of~$c$ define a \defn{corner} (resp.~a \defn{hinge}) of~$c$ if they bound a convex (resp.~concave) angle of~$c$.
%The \defn{concave chains} of~$c$ are the maximal sequences of wiggly arcs on the external boundary of~$c$ connected by hinges of~$c$.
\end{definition}

\begin{definition}
A \defn{wiggly pseudotriangle} (resp.~\defn{pseudoquadrangle}) is a cell of degree~$3$ (resp.~$4$).
\end{definition}

\begin{proposition}
\label{prop:wigglyPseudotriangulations}
Any wiggly pseudodissection~$D$ contains at most~$2n-1$ internal wiggly arcs and at most~$n$ wiggly cells.
Moreover, the following are equivalent:
\begin{compactenum}[(i)]
\item all wiggly cells of~$D$ are wiggly pseudotriangles,
\item $D$ contains~$n$ wiggly cells and is connected,
\item $D$ contains $2n-1$ internal wiggly arcs,
\item $D$ is an inclusion maximal wiggly pseudodissection.
\end{compactenum}
We then say that~$D$ is a \defn{wiggly pseudotriangulation}.
\end{proposition}

\begin{example}
\label{exm:allSmallWigglyPseudotriangulations}
For~$n = 1$, the $2$ wiggly pseudotriangulations of $3$ points are \smash{\raisebox{-.3cm}{\includegraphics[scale=.85]{wigglyPseudotriangulation1}}} and \smash{\raisebox{-.3cm}{\includegraphics[scale=.85]{wigglyPseudotriangulation2}}}.
For~${n = 2}$, the $14$ wiggly pseudotriangulations of $4$ points are \\[.2cm]
\centerline{\includegraphics[scale=.85]{wigglyPseudotriangulations}}
\end{example}

%\begin{remark}
%\label{rem:numberWigglyPseudotriangulations}
%The number~$wp_n$ of wiggly pseudotriangulations of~$n+2$ points is given by
%\[
%\begin{array}{c|ccccccccc}
%n & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & \dots \\
%\hline
%wp_n & 2 & 14 & 176 & 3232 & 78384 & 2366248 & 85534176 & 3602770400 & \dots
%\end{array}
%\]
%To compute~$wp_n$, denote by $wp_n(x)$ the polynomial where the coefficient of~$x^i$ is the number of wiggly pseudotriangulations of~$n+2$ points with~$i$ \defn{final} internal wiggly arcs (\ie ending at the last point~$n+1$).
%For instance,
%\begin{align*}
%%	wp_0(x) & = 1, \\
%	wp_1(x) & = x + 1, \\
%	wp_2(x) & = 3 x^3 + 5 x^2 + 4 x + 2, \\
%	wp_3(x) & = 15 x^5 + 35 x^4 + 44 x^3 + 40 x^2 + 28 x + 14, \\
%	wp_4(x) & = 105 x^7 + 315 x^6 + 520 x^5 + 630 x^4 + 620 x^3 + 514 x^2 + 352 x + 176.
%\end{align*}
%We invite the reader to check the expressions of~$wp_1(x)$ and~$wp_2(x)$ with \cref{exm:allSmallWigglyPseudotriangulations}.
%We use this additional variable~$x$ to obtain a recursive formula for~$wp_n(x)$.
%%Indeed, any wiggly pseudotriangulation on~$n+1$ points can be obtained from a wiggly pseudotriangulation on~$n+2$ points by collapsing the points~$n$ and~$n+1$.
%Indeed, any wiggly pseudotriangulation of~$n+2$ points can be obtained from a wiggly pseudotriangulation of~$n+1$ points by inserting a single wiggly pseudotriangle with a hinge at~$n$.
%%There are two kinds of wiggly pseudotriangulations of~$n+2$ points:
%%\begin{itemize}
%%\item those with the wiggly arc~$(n, n+1, \varnothing, \varnothing)$ are obtained from a wiggly pseudotriangulation of~$n+1$ points by choosing the position of the only corner of~$t_n$ which is not at~$n+1$,
%%\item those without the wiggly arc~$(n, n+1, \varnothing, \varnothing)$ are obtained from a wiggly pseudotriangulation of~$n+1$ points by choosing the positions of the two corners of~$t_n$ which are not at~$n+1$.
%%\end{itemize}
%Each wiggly pseudotriangulation of~$n+1$ points with $i$ final internal wiggly arcs contributes to:
%\begin{compactitem}
%\item $i+2$ wiggly pseudotriangulations of~$n+2$ points with $i+2$ final internal wiggly arcs, containing the wiggly arc~$(n, n+1, \varnothing, \varnothing)$,
%\item $j+1$ wiggly pseudotriangulations of~$n+2$ points with $j$ final internal wiggly arcs, not containing the wiggly arc~$(n, n+1, \varnothing, \varnothing)$,  for all~$0 \le j \le i+1$.
%\end{compactitem}
%\pagebreak
%This translates to the recursive formula %for all~$n \ge 1$,
%\begin{align*}
%wp_{n}(x) & = \overbracket[.5pt]{x \cdot \frac{\partial}{\partial x} \big( x^2 wp_{n-1}(x) \big)}^{\substack{\text{wiggly pseudotriangulations} \\ \text{containing } (n, n+1, \varnothing, \varnothing)}} + \overbracket[.5pt]{\frac{\partial}{\partial x}\Big( \smash{\frac{wp_{n-1}(1)-x^3wp_{n-1}(x)}{1-x}} \Big)}^{\substack{\qquad\; \text{wiggly pseudotriangulations} \qquad\;  \\ \text{not containing } (n, n+1, \varnothing, \varnothing)}} \\
%& = \frac{1}{(1 - x)^2} \big( wp_{n-1}(1) + x^2 (2 x^2 - 2 x - 1) wp_{n-1}(x) + x^4 (x - 1) wp_{n-1}’(x) \big).
%\end{align*}
%This enables us to quickly compute~$wp_n(x)$ and we obtain the numbers of wiggly pseudotriangulations~$wp_n$ by evaluating~$wp_n(x)$ at~$x = 1$.
%\end{remark}

\begin{definition}
A \defn{wiggly diagonal} of a cell~$c$ is a wiggly arc inside~$c$ and compatible with~$c$.
\end{definition}
